---
title: "Exercise-05-Quarto-Challange-1"
format: html
editor: visual
---

# Exercise 05: **Generate Sampling Distributions and CIs**

## Challenge 1

### Step 1

Loading in the packages and the "*IMDB-movies.csv*" file that I will be using for **Challange 1.** I named the file **md** instead of d.

```{r}
#| output: FALSE

library(tidyverse)
library(mosaic)
library(radiant)
mf <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/IMDB-movies.csv"
md <- read_csv(mf, col_names = TRUE)
```

### Step 2

I created a one-line statement to do some data wrangling while using the pipe operator:

-   Filtering out data that includes movies from 1920 to 1979.

-   Filtering out data that includes movies that are between 1 and 3 hours long.

-   Creating a new column that codes the startYear into a new variable **"decade"**

```{r}

md2 <- md |>
  filter(startYear >= 1920 & startYear <= 1979) |>
  filter(runtimeMinutes >= 60 & runtimeMinutes <= 180) |> 
  mutate(decade = case_when(startYear >= 1920 & startYear <= 1929 ~ "20s",
                            startYear >= 1930 & startYear <= 1939 ~ "30s",
                            startYear >= 1940 & startYear <= 1949 ~ "40s",
                            startYear >= 1950 & startYear <= 1959 ~ "50s",
                            startYear >= 1960 & startYear <= 1969 ~ "60s",
                            startYear >= 1970 & startYear <= 1979 ~ "70s"))
md2
```

### Step 3

Using {ggplot} I plotted histograms of the distribution of **"runtimeMinutes"** for each **"decade."** I was not sure what our y-axis should be so I left it as counts/frequency (vs density).

```{r}
#| warning: FALSE 

md2_p <- ggplot(md2, aes(x = runtimeMinutes)) + 
  geom_histogram(color="black", fill="orange") +
  facet_wrap(~decade, ncol = 3) +
  labs(x = "Runtime (Minutes)", y = "Counts", title = "Distribution of Runtime by Decades")

md2_p
```

### Step 4

I created a one-line statement to calculate the:

-   population mean

-   population standard deviation (SD)

This was done for **"runtimeMinutes"** for each **"decade"** and then was saved in a new dataframe called **population_results.** I used the function **sdpop()** function from the ***{radiant}*** package to calculate the standard deviation (SD) for the population.

```{r}

population_results <- md2 |> 
  group_by(decade) |>
  summarize(meanRTMpopulation = mean(runtimeMinutes, na.rm = TRUE), 
            sdRTMpopulation = sdpop(runtimeMinutes, na.rm = TRUE))
population_results
```

### Step 5

I single sampled 100 movies without replacement from the original data set and for each decade using the **slice_sample()** function. I then calculated the:

-   single sample mean

-   single sample standard deviation

This was done for the **"runtimeMinutes"** for each **"decade."** I also created a new dataframe called **mdsample.** Values for the mean and standard deviation are always changing every time you run the code since it's taking a new sample. So, I ended up splitting the mode into two chunks so I can work with the same numbers. I couldn't set the **set.seed() function** to work.

```{r}

md_sample <- md2 |> 
  group_by(decade) |>
  slice_sample(n = 100, replace = FALSE)
md_sample
  
md_sampleMSD <- md_sample |> 
  group_by(decade) |>
  summarize(meanRTMsample = mean(runtimeMinutes, na.rm = TRUE), 
            sdRTMsample = sd(runtimeMinutes, na.rm = TRUE))
md_sampleMSD
```

### Step 6

I calculated the estimated standard error (SE) for the single sample of **"runtimeMinutes"** I took above in Step 5. This was based on the standard deviation and the sample size (n=100).

```{r}

md_sampleSEM <- md_sample |>
  group_by(decade) |>
  summarize(semRTMsample = sd(runtimeMinutes, na.rm = TRUE) / sqrt(length(runtimeMinutes)))

md_sampleSEM

```

### **Step 7**

I first calculated the standard error (SE) for the entire population.

```{r}
md_populationSEM <- md2 |>
  group_by(decade) |> 
  summarize(semRTMpopulation = sd(runtimeMinutes, na.rm = TRUE)/sqrt(length(runtimeMinutes)))

md_populationSEM

```

I then created a new table using the **innerjoin()** function so we could compare the means, standard deviation (SD), and the standard error (SE) of the sample we took and the entire population.

```{r}
sample_population_comparison <- population_results |> 
  inner_join(md_populationSEM, by = "decade") |>
  inner_join(md_sampleMSD, by = "decade") |>
  inner_join(md_sampleSEM, by = "decade")
sample_population_comparison

```

### **Step 8**

I attempted to create a *sampling distribution* of mean **"runtimeMinutes"** for each **"decade":**

-   I drew 1000 random samples of 100 movies from each decade, without replacement

-   Then for each sample I calculated the mean and standard deviation (SD) for **"runtimeMinutes"** for each decade.

    ```{r}

    random_sample <- 1000
    sample_size <- 100
    s <- list() #dummy variable to hold the samples

    for (i in 1:random_sample) {
       s[[i]] <- md2 |> #md2 original table with decades column
         group_by(decade) |> 
         slice_sample(n = sample_size, replace = FALSE) |> # used slice_sample for datatables 
         group_by(decade) |>
         summarize(meanRTMsampledis = mean(runtimeMinutes, na.rm = TRUE), 
                   sdRTMsampledis = sd(runtimeMinutes, na.rm = TRUE))
    }

    combined_sample_distribution  <- bind_rows(s) #combined everything into one datatable
    combined_sample_distribution  
    ```

### **Step 9**

I then calculated the mean and and the standard deviation (SD) of the sampling distribution of the sample means for each **"decade."**

-   mean: should be a good estimate of the population mean

-   standard deviation (SD): estimate of the standard error of the population

    ```{r}
    sample_distributionMSD <- combined_sample_distribution|> # using the combined table above
      group_by(decade) |> 
      summarize(meanRTMcombinedsampledis = mean(meanRTMsampledis, na.rm = TRUE), 
                sdRTMcombinedsampledis = sd(meanRTMsampledis, na.rm = TRUE))
    sample_distributionMSD
    ```

I then plotted a histogram of the sampling distribution for each **"decade."** Now each **"decade"** has normal distribution bell shape.

```{r}
#| warning: FALSE 

combined_sample_distribution_p <- 
  ggplot(combined_sample_distribution, aes(x = meanRTMsampledis)) + 
  geom_histogram(color="black", fill="purple") +
  facet_wrap(~decade, ncol = 3) +
  labs(x = "Runtime (Minutes)", y = "Counts", title = " Sampling Distribution of Runtime by Decades")

combined_sample_distribution_p
```

### **Step 10**

I compared the standard errors (SE) **"runtimeMinutes"** for each **"decade"** for:

-   SE for the first sample of 100 movies

-   SE for the known population

-   SE for the sampling distribution of sample means

```{r}

#this first part is just creating the SE for the sampling distribution
sample_distributionSEM <- combined_sample_distribution|>
  group_by(decade) |> 
  summarize(sdRTMcombinedsampledis = sd(meanRTMsampledis, na.rm = TRUE))


standard_error_comparison <- md_sampleSEM |>
  inner_join(md_populationSEM, by = "decade") |>
  inner_join(sample_distributionSEM, by = "decade")

standard_error_comparison
```
